\documentclass[a4paper, 12pt]{article}
\usepackage[margin=2cm]{geometry} 
\usepackage{amsmath,amsfonts}
\usepackage{amssymb,amsthm}
\usepackage{tikz,tkz-euclide}
\usepackage[cm]{fullpage}
\usepackage{fancyvrb}

\usetikzlibrary{calc,patterns,angles,quotes}
\usetkzobj{all}

\title{Senior Test 4}
\author{Stellenbosch Camp 2018}
\date{Time: $2 \frac{1}{2}$ hours}

\begin{document} \maketitle

\begin{enumerate}
% 2017/2018 British MO Round 2 Q1
\item[1.]Assume that $a_1,a_2,a_3,...$ is an infinite strictly increasing sequence of positive integers, and $p_1,p_2,p_3,...$ is a sequence of primes such that $p_n | a_n$ for every positive integer $n$ and $k$. Prove that all the numbers $a_1,a_2,...$ are primes.
\vspace{7pt}

\item[2.]A polynomial $P(x)$ is chosen so that each of the polynomials $P(P(x)) $ and $P(P(P(x))) $ is strictly monotone on the real axis. Prove that $P(x)$ is also strictly monotone on the real axis.

\item[3.]Two arbitrary positive integers $a$ and $b$ are given. Prove that there exists infinitely many positive integers $n$ such that $n^b +1 \nmid a^n +1$

\vspace{5pt}

\item[4.]A positive integer $k$ is given. Initially, $N$ cells are marked on an infinite chequered plane. We say that the \textit{cross} of a cell is the set of all cells lying in the same row or in the same column as $A$. By a turn, it is allowed to mark an unmarked cell $A$ if the cross of $A$ already contains at least $k$ marked cells. It appears that every cell can be marked in a stretch of turns. Determine the smallest possible value of $N$
\vspace{7pt}

\item[5.]Points $P$ and $Q$ are chosen respectively on the sides $AB$ and $AC$ of a triangle $ABC$ so that $PQ||BC$. The segments $BQ$ and $CP$ meet at $O$. Let $A'$ be the reflection of $A$ with respect to the line $BC$. The segment $A'O$ meets the circumcircle $\omega$ of the triangle $APQ$ at $S$. Prove that the circumcircle of the triangle $BSC$ is tangent to $\omega$.
\vspace{7pt}


\end{enumerate}
\end{document}
