\documentclass{article}

\usepackage{amsmath,amsthm,amssymb}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{hyperref}

\DeclareMathOperator{\lcm}{lcm}


\begin{document}

\begin{center}
\textbf{\Large Junior February Monthly Problem Set}
\\ \vspace{1em}
\textbf{\large Due: 22 February 2019}
\end{center}

\vspace{12pt}

\begin{enumerate}[1.]

\item % IPMO 2006 Junior Individual Q5
If $a = 2^{2006} + 2^{-2006}$ and $b = 2^{2006} - 2^{-2006}$ then what is the value of $a^2 - b^2$?


\item % IPMO 2008 Junior Individual Q11
Paul likes dogs. At present all his adult dogs are Dalmatians while some of his puppies are Dalmatians and some are not. In all he has 11 dogs of which 7 are Dalmatians and 8 are puppies. How many Dalmatian puppies has he?


\item % IPMO 2006 Junior Team Q6
How many distinct factors does 1 000 000 000 have?


\item % IPMO 2008 Junior Team Q9
If a regular octagon has side length 1 cm, find the distance between the opposite sides.


\item % SW-2013-3
Let $n$ be a positive integer. Both $n$ and $n^2$ only contain the digits $1$, $2$ and $3$ (not necessarily all of them). Determine all possible values of $n$.


\item % DB-2012-2
Given a (not necessarily convex) quadrilateral $ABCD$, call a point $P$ in the same plane as $ABCD$ an \emph{areal centre} for $ABCD$, if any line through $P$ divides $ABCD$ into two parts of equal area. What are necessary and sufficient conditions on $ABCD$ for it to possess an areal centre?


\item % The Liam
The (English language version of the) game of Scrabble\texttrademark consists of 100 tiles, each containing either a letter from A to Z (some letters occur more than once), except for two blank tiles; see \href{https://en.wikipedia.org/wiki/Scrabble_letter_distributions#English}{\texttt{the relevant Wikipedia page}} for the exact distribution of multiplicities of each letter.

In a solo game of Scrabble, the player starts by choosing seven tiles from the 100 available tiles at random. What is the probability that the player picks up exactly two vowels?


\item % 


\end{enumerate}


\vfill
\textbf{\Large Email submission guidelines}
\begin{itemize}
	\item Email your solutions to \href{mailto:samf.training.assignments@gmail.com}{\texttt{samf.training.assignments@gmail.com}}.
	\item Submit each question in a single separate PDF file (with multiple pages if necessary), with your name and the question number written on each page.
	\item If you take photographs of your work, use a document scanner such as CamScanner to convert to PDF.
	\item If you have multiple PDF files for a question, combine them using software such as PDFsam.
\end{itemize}

\end{document}
